Ab Initio Effective One-Electron Potentials Revisited: General Theory and Example Applications for Charge-Transfer Energy Calculations

### 2019-11-05 papers/2_oep_technique/paper-r1.tex
Paper II revision

\section{\label{s:1}Introduction}

\section{\label{s:2}Effective One-Electron Potentials: General Theory}

The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. 
For convenience, the total effective density can be split into nuclear and electronic contributions,
%
\begin{equation} \label{e:rho-eff}
 \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
\end{equation}
%
where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
Formally, the electronic part can be expanded in terms of an effective
one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = 
        \lambda \hat{v}_{\rm nuc} +
        %\sum_{\alpha\beta} 
	%\Ket{\alpha} 
        \int d{\bf r} \Ket{{\bf r}} 
        v^{\rm eff}({\bf r})
	%\left[
	%\int
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
	%\rho^{\rm eff}({\bf r}')
	%d{\bf r}' 
	%\right]
	%\Bra{\beta}  
        \Bra{{\bf r}}
\end{equation}
%
with the nuclear
%and Coulomb operators 
operator
defined by
%
%\begin{subequations}
%\begin{align}
\begin{equation}
        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
                     \int d{\bf r} \Ket{{\bf r}} 
                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
                     \Bra{{\bf r}} \;.
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
	%f({\bf r}) 
	%\int
	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%\end{align}
%\end{subequations}
%
%for any one\hyp{}electron function $f({\bf r})$. 
and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
The matrix element of the effective potential operator
is therefore given by
%
\begin{equation}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;,
\end{equation}
%
where 
%
\begin{equation}
 W_{\alpha\beta}^{(x)} = 
 Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
\end{equation}
%
$Z_x$ is the atomic number of the $x$th atom,
and the symbol `At' denotes all atoms that contribute to the effective potential.
In the above equation and throughout the work, 
%
\begin{equation}
\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathscr{O}(1)$ 
and the electron repulsion integral (ERI)
is defined according to
%
\begin{equation}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

\section{\label{s:3}Effective One-Electron Potentials: Configuration Interaction}

The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
electronic state
of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
%
\begin{equation} \label{e:schrodinger}
 E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
\end{equation}
%
for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator,
$\hat{h}(1) = \hat{v}_{\rm nuc} -\frac{1}{2}\sum_e^{\rm el} \nabla_e^2$,
and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
From the above solution, any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation} \label{e:prop}
 P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
\end{equation}
%
The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
the configuration interaction (CI) expansion around the reference 
(approximate) wavefunction, $\Psi^{(0)}$,
typically the solution to the Hartree\hyp{}Fock (HF) problem\cite{Roothaan.RevModPhys.1951},
%
\begin{equation} \label{e:ci}
 \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
\end{equation}
%
In the above equation,
$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
virtual (unoccupied) molecular orbital, obtained from the HF solution.
By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
operator matrix element between
the CI configurations, it can be shown that the following holds
%
\begin{equation} \label{e:exp-val-series}
	P =
	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
\end{equation}
%
%where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $
%$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
%%
%\begin{equation}
%	\BraKet{ij}{kl} \equiv
%	\iint 
%	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
%	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
%	d{\bf r}_1 d{\bf r}_2  \;,
%\end{equation}
%%
where the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
are well defined and characteristic for the property of interest.
In fact, they are certain functions of the CI coefficients, i.e.,
%
\begin{equation}
{\bf P} = {\bf P}
\left(
 \left\{
  C_{r}^{a}, C_{rs}^{ab}, \ldots 
 \right\} 
\right)
\end{equation}
%
and, in principle, are known as long as the solution to the CI problem 
or its approximation is available.

%In this work we are interested in cases when

%$\Psi^{AB}$. In this Section we consider theories to calculate the effect of the interaction
%of molecules $A$ and $B$ on the property $P$ knowing only the unperturbed wavefunctions,
%$\Psi^{A}$ and $\Psi^{B}$.

\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Property}

Consider a molecular aggregate consisting of two interacting 
molecules $A$ and $B$. The part of property that arises due to the
intermolecular interaction between $A$ and $B$ can be calculated
based on the supermolecular approach,
%
\begin{multline} \label{e:delta-p}
	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} \\- 
	\left(
	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} +
	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}
	\right)
	\;,
\end{multline}
%
where $\Psi^{AB}$ and $\Psi^{A(B)}$ are the wavefunctions of the interacting molecular aggregate
and the unperturbed (non\hyp{}interacting) molecules, respectively.
The notation $\Delta^{AB}$ emphasizes on the two\hyp{}body character of the interaction\hyp{}induced
property. In this work we focus on re\hyp{}expressing $\Delta^{AB}P$ in terms of separate
fragments associated with either molecule. In general case when the basis set
is complete, such an operation is not possible. However,
assuming an approximation
in which the basis functions $\phi_i$ 
are localized on either molecule it allows one to partition the basis set space
and the resulting 2\hyp{} and 4\hyp{}index tensor elements
into subsets `belonging' either to molecule $A$ or to molecule $B$.
Such an approximation is essentially a foundation of any fragment\hyp{}based strategy
and, in particular, designing \emph{ab initio} force fields such as EFP2
of SolEFP.

Application of Eq.~\eqref{e:exp-val-series}
to Eq.~\eqref{e:delta-p} and adopting the localized basis set approximation
yields the partitioning of the interaction\hyp{}induced property,
%
\begin{equation} \label{e:delta-p-part}
 \Delta^{AB} P = \Delta^{AB} P(1) + \Delta^{AB} P(2) \;,
\end{equation}
%
for which the one\hyp{}electron part reads
%
\begin{multline} \label{e:delta-p-part-1el}
 \Delta^{AB} P(1) \approx
	\sum_{ij} \Big\{ 
	P^{AA}_{ij}  \tBraKet{A}{\hat{h}}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
	P^{BB}_{ij}  \tBraKet{B}{\hat{h}}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
2	P^{AB}_{ij}  \tBraKet{A}{\hat{h}}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
	\Big\}
\end{multline}
%
and the two\hyp{}electron part is given by
%
\begin{multline} \label{e:delta-p-part-2el}
 \Delta^{AB} P(2) \approx
	\sum_{ijkl} \Big\{ 
	P^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
	P^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
	{s} \left[ P^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
	{s} \left[ P^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
	\Big\} \;.
\end{multline}
%
To simplify the resulting expressions
in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
the shorthand notation $\tBraKet{A}{\hat{h}}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
was introduced, and
the real orbitals 
were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
The operator $s$ permutes and sums over all the combinations of the basis function indices
given their partitioning scheme between molecules $A$ and $B$.
All the summation components from Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el} 
can be grouped into the three
sub\hyp{}categories: (i) Coulomb\hyp{}like terms
with integrals of the type $\BraKet{AA}{BB}$, (ii) overlap\hyp{}like terms with $\BraKet{AA}{AB}$
and $\BraKet{AB}{BB}$ and (iii) exchange\hyp{}like terms with $\BraKet{AB}{AB}$.

%Therefore, the interaction\hyp{}induced property can be recast in a following way

@@

\subsection{\label{ss:2.4}Definition of Interaction-Induced Property Effective Potentials}

The above technique can be now applied to the CI expansion of the interaction\hyp{}induced property
under the localized basis set approximation 
from Eq.~\eqref{e:delta-p-part}.
%Coulomb\hyp{}like and overlap\hyp{}like 
%terms from Eq.~\eqref{e:delta-p-part-2el}. 
For example, the Coulomb\hyp{}like contributions
can be rewritten without making any further approximation as
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
 \equiv
	\sum_{kl\in B} 
	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
with
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
\end{equation}
%
Similarly for the overlap\hyp{}like term we have
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
 \equiv
	\sum_{k\in A} \sum_{l\in B} 
	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
and the associated effective potential reads
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
\end{equation}
%
Note that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
%
\begin{multline} \label{e:ci-oep-master}
 \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
	+ \sum_{i\in A} \sum_{j\in B} 
	  \tBraKet{i}{ \left\{ \hat{v}^{{\rm eff},AB}_{1(ij)} + \hat{v}^{{\rm eff},BA}_{1(ij)}  \right\} }{j}
	 \\ +
	\sum_{i\in A} \sum_{k\in A}
	  \tBraKet{i}{\hat{v}^{{\rm eff},BA}_{2(ik)} }{k}
	 +
        \sum_{j\in B} \sum_{l\in B}
          \tBraKet{j}{\hat{v}^{{\rm eff},AB}_{2(jl)} }{l}
\end{multline}
%
where the exchange\hyp{}like component is
%
\begin{equation}
	\Delta_{\rm ex}^{AB} P = 
	s \sum_{ik\in A} \sum_{jl\in B}
	P^{ABAB}_{ijkl} \BraKet{ij}{kl} \;,
\end{equation}
%
and
the overlap\hyp{}like and the Coulomb\hyp{}like 
OEP's are defined as follows
%
\begin{subequations}
\begin{align} \label{e:oep-exch-like}
	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \;, \\
	%
	\hat{v}^{{\rm eff},BA}_{2(ik)} &\equiv P_{ik}^{AA} \hat{h} 
        + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
        \\ \label{e:oep-coul-like}
	& \qquad\qquad+ \frac{1}{2}
        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} \;,
\end{align}
\end{subequations}
%
and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.

\subsection{\label{s.242}Separability of P-Coefficients}

Until that point, the effective potentials from Eqs.~\eqref{e:oep-exch-like}
and \eqref{e:oep-coul-like} are not separable in terms of the interacting molecules
because the $\bf P$ tensors are derived from the CI wavefunction of
the entire interacting system $AB$. However, if there exist an approximate
separation as, for example,
%
\begin{align}
 P^{AAAB}_{mkij} &\approx X^A_{mk} Y^{AB}_{ij} \;, \\
 P^{AABB}_{ikjl} &\approx Z^A_{ik} Y^{AB}_{} Z^{ B}_{jl} \;,
\end{align}
%
and if the matrices ${\bf X}^{A(B)}$, 
${\bf Z}^{A(B)}$ and ${\bf P}^{AA(BB)}$ as well as the tensors
${\bf P}^{AAAA(BBBB)}$ 
are properties of the unperturbed (isolated) molecules,
further rearrangements in Eq.~\eqref{e:ci-oep-master} are possible
leading to computationally more efficient expressions.

Note also that the matrices ${\bf Y}^{AB}$ and ${\bf P}^{AB}$ depend on their instantaneous 
relative positions.

%then the molecule\hyp{}specific OEP's can be defined.

\section{\label{s.334}Practical OEP-Based Calculations}

There are two general cases of basis function partitioning scheme
in which it is possible to define an effective one\hyp{}electron
potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{A}_{\rm eff}}{j} $ && $\tBraKet{j}{\hat{v}^{A}_{\rm eff}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[2]/RI\footnotemark[3] Form    
&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[4] Form                     
&& --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Test footnotemark.}
\footnotetext[2]{Density Fitting}
\footnotetext[3]{Resolution of Identity}
\footnotetext[4]{Distributed Multipole Expansion}
%
\end{table}
}
%

@@

This scheme is most applicable in the case of matrix elements of the type
$
 \tBraKet{j^B}{\hat{v}_{\rm eff}^A}{l^B}
$
because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
and $\rho^A_{\rm eff}$. 
In general, given certain two effective one\hyp{}electron density distributions,
the associated effective Coulombic interaction energy can be estimated from the classical formula
according to
%
\begin{equation}
 \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
 {\vert {\bf r}_1 - {\bf r}_2 \vert} 
d{\bf r}_1 d{\bf r}_2 \;.
\end{equation}
%
The above double integral can be approximated by the
distributed multipole (DMTP)
expansions of the respective densities as follows
%
\begin{multline}
  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
 \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
 \sum_{u\in A} \sum_{w\in B} \Big\{ 
 T^{(0)}_{uw}
 q_{\rm eff}^{(u)}  q_{\rm eff}^{(w)} \\
 - {\bf T}^{(1)}_{uw} \cdot 
   \left[ q_{\rm eff}^{(u)} {\BM\upmu}_{\rm eff}^{(w)} - q_{\rm eff}^{(w)} {\BM\upmu}_{\rm eff}^{(u)} \right]
 - {\bf T}^{(2)}_{uw} : 
  {\BM\upmu}_{\rm eff}^{(u)}  \otimes {\BM\upmu}_{\rm eff}^{(w)} 
 \ldots
 \Big\}
\end{multline}
%
In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
are the effective distributed monopole (charge)
and dipole moment, respectively, centered on the $u$th site, 
whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
that can be found elsewhere.
The symbol `$\odot$' denotes the sum of all the tensor contractions
performed over the DMTP's of molecule $X$ and $Y$ to yield the associated interaction energy.
The choice of the distribution centres as well as the truncation order of the multipole expansion
is crucial in compromising the accuracy and computational cost of the resulting expressions.
There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
of Stone and Anderton, %cite
the cumulative atomic multipole moments of Sokalski and Poirier, %cite
the localised distributed multipole expansion (LMTP) of Etchtebest et al. %cite
or schemes based on fitting to electrostatic potential.

@@

\section{\label{s:5}Results and Discussion}
\subsection{\label{ss.5.1}Pauli-Repulsion Interaction Energy}

At short intermolecular interactions, wavefunction overlap increases
which causes the electrons from either molecules to move apart to minimize the repulsive interactions.
The resulting interaction energy can be estimated based on perturbation theory approaches.
In this section, we demonstrate the use of OEP operators to significanlty reduce the computational
cost of such calculations, that can be used in fragment\hyp{}based modelling of extended molecular
aggregates in terms of force field methods.

>>>>

%\section{\label{s:1}Introduction}

\section{\label{ss.5.2}Introduction}

Charge transfer (CT) between molecules occurs when the net electronic populations 
of interacting molecules change which leads to an additional stabilization 
of a molecular aggregate. The effect of CT on the interaction energy 
is known to be often non\hyp{}negligible, especially
in donor\hyp{}acceptor systems such as H\hyp{}bonded species
and charged complexes.
\emph{Ab initio} fragmentation methods, such as
the second generation of the effective fragment potential method (EFP2),
are often used to describe the homogeneous extended molecular systems 
such as solutions
and recently even biomolecules. EFP2 relies on the intermolecular 
perturbation theory, and the total interaction energy is approximated as
a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
charge densities of the monomers, exchange\hyp{}repulsion energy,
induction energy due to mutual polarization, intermolecular dispersion energy,
as well as the CT energy,
%
\begin{equation}
 U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
\end{equation}
%
However, calculating the CT stabilization energy is usually much more expensive 
than evaluating the contributions due to other effects.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
%However, the computational cost of the CT term is usually at least few times larger
%than the next costly contribution $U^{\rm Ex-Rep}$, which is still too costly
%to perform molecular dynamics (MD) simulations including the $U^{\rm CT}$. 
In effect,
this part of the EFP2 energy is usually ignored in most of the applications. In the 
recent LIBEFP library, CT term were even not implemented due to those reasons.

In this work, we propose an alternative formulation of the CT energy compatible with the
EFP2 method. We start from the perturbation theory of Murrell et al.
and
introduce the effective one\hyp{}electron potential (OEP) operators
to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
Later we validate the model against the EFP2 method in Section XXX. Finally,
we conclude this work and provide few remarks on future development in Section XXX.

\section{\label{s:2}Effective One-Electron Potential Operators}

To establish the notation within the present work, 
the formalism of one\hyp{}electron potential (OEP) operators is
given first.
The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. 
For convenience, the total effective density can be split into nuclear and electronic contributions,
%
\begin{equation} \label{e:rho-eff}
 \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
\end{equation}
%
where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
Formally, the electronic part can be expanded in terms of an effective
one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = 
        \lambda \hat{v}_{\rm nuc} +
        %\sum_{\alpha\beta} 
	%\Ket{\alpha} 
        \int d{\bf r} \Ket{{\bf r}} 
        v^{\rm eff}({\bf r})
	%\left[
	%\int
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
	%\rho^{\rm eff}({\bf r}')
	%d{\bf r}' 
	%\right]
	%\Bra{\beta}  
        \Bra{{\bf r}}
\end{equation}
%
with the nuclear
%and Coulomb operators 
operator
defined by
%
%\begin{subequations}
%\begin{align}
\begin{equation}
        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
                     \int d{\bf r} \Ket{{\bf r}} 
                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
                     \Bra{{\bf r}} \;.
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
	%f({\bf r}) 
	%\int
	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%\end{align}
%\end{subequations}
%
%for any one\hyp{}electron function $f({\bf r})$. 
and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
The matrix element of the effective potential operator
is therefore given by
%
\begin{equation}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;,
\end{equation}
%
where 
%
\begin{equation}
 W_{\alpha\beta}^{(x)} = 
 Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
\end{equation}
%
$Z_x$ is the atomic number of the $x$th atom,
and the symbol `At' denotes all atoms that contribute to the effective potential.
In the above equation and throughout the work, 
%
\begin{equation}
\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathscr{O}(1)$ 
and the electron repulsion integral (ERI)
is defined according to
%
\begin{equation}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

%\subsection{\label{s.334}Practical OEP-Based Calculations}

@@

Two general cases of basis function partitioning scheme
are considered in this work,
in which it is possible to define an effective one\hyp{}electron
potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{A}_{\rm eff}}{j} $ && $\tBraKet{j}{\hat{v}^{A}_{\rm eff}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[2]/RI\footnotemark[3] Form    
&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[4] Form                     
&& --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Test footnotemark.}
\footnotetext[2]{Density Fitting}
\footnotetext[3]{Resolution of Identity}
\footnotetext[4]{Distributed Multipole Expansion}
%
\end{table}
}
%

@@

This scheme is most applicable in the case of matrix elements of the type
$
 \tBraKet{j^B}{\hat{v}_{\rm eff}^A}{l^B}
$
because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
and $\rho^A_{\rm eff}$. 
In general, given certain two effective one\hyp{}electron density distributions,
the associated effective Coulombic interaction energy can be estimated from the classical formula
according to
%
\begin{equation}
 \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
 {\vert {\bf r}_1 - {\bf r}_2 \vert} 
d{\bf r}_1 d{\bf r}_2 \;.
\end{equation}
%
The above integral can be approximated by the
distributed multipole (DMTP)
expansions of the respective densities as follows
%
\begin{multline}
  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
 \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
 \sum_{u\in A} \sum_{w\in B} \Big\{ 
 T^{(0)}_{uw}
 q_{\rm eff}^{(u)}  q_{\rm eff}^{(w)} \\
 - {\bf T}^{(1)}_{uw} \cdot 
   \left[ q_{\rm eff}^{(u)} {\BM\upmu}_{\rm eff}^{(w)} - q_{\rm eff}^{(w)} {\BM\upmu}_{\rm eff}^{(u)} \right]
 - {\bf T}^{(2)}_{uw} : 
  {\BM\upmu}_{\rm eff}^{(u)}  \otimes {\BM\upmu}_{\rm eff}^{(w)} 
 \ldots
 \Big\}
\end{multline}
%
In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
are the effective distributed monopole (charge)
and dipole moment, respectively, centered on the $u$th site, 
whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
that can be found elsewhere.
The symbol `$\odot$' denotes the sum of all the tensor contractions
performed over the DMTP's of molecule $X$ and $Y$ to yield the associated interaction energy.
The choice of the distribution centres as well as the truncation order of the multipole expansion
is crucial in compromising the accuracy and computational cost of the resulting expressions.
There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
of Stone and Anderton, %cite
the cumulative atomic multipole moments of Sokalski and Poirier, %cite
the localised distributed multipole expansion (LMTP) of Etchtebest et al. %cite
or schemes based on fitting to electrostatic potential.

@@

\section{\label{s:34r}Charge Transfer Interaction Energy for Fragment Potentials}

\subsection{\label{s:34rg} Murrell et al.'s Model: Starting Point}
